---
title: "State Regulation of Social Media and Children in the United States"
author:
  - name: Oliver Huang
    affiliations:
      - SPRING Program on Health and Society
      - Duke University
  - name: Khadija Mohammad
    affiliations:
      - SPRING Program on Health and Society
  - name: Arik Karim
    affiliations:
      - SPRING Program on Health and Society
      - The George Washington University
format: pdf
editor: visual
execute: 
  echo: false
  output: true
bibliography: references.yaml
---

# Abstract

Over the past year, political sentiment has pushed states to regulate social media use for children in the United State. We analyze 300 bills introduced by US state legislators in 2024. We find that while Democratic- and Republican-controlled legislators both introduced legislation related to social media, Republican-controlled state legislatures were more likely to enact such regulation.

```{r}
#| label: add-library
#| echo: false
#| output: false
library(tidyverse)
library(usmap)
library(patchwork)
library(knitr)
```

# Methods

We scraped the data from the National Conference of State Legislatures [@socialm]. The data was imported on October 3, 2024. We utilized rstudio, along with the tidyverse package and usmap package to read in and analyze the data.

In total, there are 50 states and 9 different statuses for the legislation, including "Enacted", "To governor", "Adopted", "Pending", "Pending - Carryover", "Failed - Adjourned", "Failed", "Vetoed", and " " for NA values. We then cleaned the data to remove all values in the dataset that indicated the state had no legislation related to social media and children.

When analyzing the data, we looked to understand which states introduced the highest number of bills. We also calculated how many bills were passed by state legislatures and in which state. Finally, we read in data from the National Conference of State Legislatures on the partisan control of state legislatures and government. We used this data to understand the political control of the state governments that introduced and enacted legislation regulating social media.

```{r}
#| label: data-read-in
#| output: false

state_leg_comp_24 <- read_csv("data/state-leg-comp-24.csv")
state_leg_comp_24 <- state_leg_comp_24 |> 
  mutate(STATE = as.factor(STATE)) |> 
  rename(state = "STATE")

state_24 <- read_csv("data/state-sm-24") 

state_24 |> 
  distinct(jurisdiction) |> 
  pull(jurisdiction)

state_24 |> 
  distinct(status) |> 
  pull(status)

state_24 <- state_24 |> 
  mutate(
    jurisdiction = as_factor(jurisdiction),
    status = as_factor(status), 
    state = jurisdiction
  )
```

```{r}
#| label: data-clean
#| output: false
state_none <- state_24 |> 
  filter(number != "None",
         number != "No regular 2024 session")

state_24 <- state_24 |> 
  filter(number != "None",
         number != "No regular 2024 session") |> 
  mutate(
    status = fct_relevel(
      status, "Enacted", "To governor", "Adopted", "Pending", "Pending - Carryover", "Failed - Adjourned", "Failed", "Vetoed", 
    )
  )

levels(state_24$status)
```

# Results

```{r}
#| label: tbl-legislation-status
#| tbl-cap: "States that passed the most legislation"
#| output: false
state_count <- state_24 |> 
  count(state, sort = "TRUE") |> 
  left_join(state_leg_comp_24) |> 
  select(state, n, Leg_Control) |> 
  write_csv("results/state-count-24.csv")
state_count

state_passed <- state_24 |> 
  filter(status == "Enacted" ) |> 
  count(state, sort = "TRUE") |> 
  left_join(state_leg_comp_24) |> 
  select(state, n, Leg_Control) |> 
  write_csv("results/state-passed.csv")
state_passed |> 
  kable()
```

```{r}
#| label: tbl-state-status
#| tbl-cap: "Status of State Bills on Social Media"
state_24 |> 
  count(status) |> 
  rename(Status = status) |> 
  rename("Number of Bills" = n) |> 
  kable()

```

We found 42 states had introduced 300 pieces of legislation related to social media and children. Out of the 42 states, 19 states ended up enacting a total of 29 pieces of legislation related to social media and children. At most, states only passed 3 bills related to social media.

```{r}
#| label: fig-bill-reg-vis
#| fig-cap: "State's with Bills Regulating Social Media in 2024"
plot_usmap(data = state_count, values = "n", color = "#333333") + 
  scale_fill_continuous(low = "white", high = "#416555", name = "Number of Bills") + 
  labs(title = "State's with Bills Regulating Social Media in 2024",
       subtitle = "States with No Bills have a Gray Fill",
       caption = "SOURCE: National Conference of State Legislatures") +
  theme(legend.position = "right")
```

```{r}
#| label: fig-state-pass-vis
#| fig-cap: "State's that Passed Bills Regulating Social Media in 2024"

plot_usmap(data = state_passed, values = "n", color = "#333333") + 
  scale_fill_continuous(low = "white", high = "#416555", name = "Number of Bills") + 
  labs(title = "State's that Passed Bills on Social Media",
       subtitle = "States with No Bills have a Gray Fill",
       caption = "SOURCE: National Conference of State Legislatures") +
  theme(legend.position = "right")
```

## Partisan Breakdown

Overall, 14 of the state legislatures are Democrat-controlled, 25 are Republican-controlled, and 3 are divided. This suggests that interest in social media regulation is relatively bipartisan across party lines.

```{r}
#| label: state-count-party
sc <- state_count |> 
  mutate(
    Leg_Control = case_when(
      Leg_Control == "Unicameral" ~ "Rep",
      .default = as.character(Leg_Control)
    )
  ) |> 
  count(Leg_Control)
```

The states that have introduced the most legislation are large Democrat or swing states, including New York, California, Illinois, New Jersey, and Virginia.

```{r}
#| label: tbl-print-most-bills
#| tbl-cap: "Partisan Control and "
head(state_count, n = 10) |> 
  rename(
    State = state,
    "Number of Bills" = n,
    "Partisan Control of Legislature" = Leg_Control
  ) |> 
  kable()
```

For states that passed legislation, the partisan control of state legislatures is mixed, with Republican controlled legislatures, such as Utah, and Democrat controlled legislatures, such as Maryland, passing bills related to social media regulation. However, only five of the 19 states that have passed legislation regarding children and social media--Colorado, Maryland, Virginia, Kentucky, and Minnesota--have Democrats controlling at least one house of the state legislature.

```{r}
#| label: print-most-bills-passed
sp_count <- state_passed |> 
  mutate(
    Leg_Control = case_when(
      Leg_Control == "Unicameral" ~ "Rep",
      .default = as.character(Leg_Control)
    )
  ) |> 
  count(Leg_Control)
```

# Discussion

Overall, it is clear that regardless of partisan affiliation, Democrat- and Republican-controlled state legislatures seem equally interested in introducing legislation regarding social media and children.

It is worth noting that while Democratic states introduced the most bills en masse, a higher proportion of Republican-controlled state legislatures passed bills.

```{r}
#| label: fig-prop-graph-partisan-leg
sc <- sc |> 
  mutate(
    type = "Introduced Legislation"
  )

sp_count <- sp_count |> 
  mutate(
    type = "Passed Legislation"
  )

sc |> 
  bind_rows(sp_count) |> 
  rename(leg_control = Leg_Control) |> 
  mutate(
    leg_control = as.factor(leg_control),
    leg_control = fct_relevel(
    leg_control, "Rep", "Divided", "Dem"
  )) |> 
  ggplot(aes(x = type, y = n, fill = leg_control)) +
  geom_col(
    position = "fill"
  ) +
  labs(
    x = " ",
    y = "Proportion",
    fill = "Legislative Control",
    title = "Partisan Control of State Legislature"
  ) +
  theme_minimal()
```

A plausible explanation for this is that while Democrats are more inclined to regulate the technology industry, there is a heavy technology lobbying presence in the state, making it harder to pass bills. Both New York and California are known for their technology industry, whose presence may lead to more lobbying in the states against bills.

## Large States and Small States

With the exception of Florida, most states that have passed legislation regulating social media have medium to low populations. This suggests two possibilities. First, large states may have larger state legislators, which leads to more deliberation, halting the passage of legislation. Second, technology companies may also be more inclined to intensely lobby in states with large populations. A larger state means a larger user base will be affected by new regulations. As a result, companies may put more effort into stalling the passage of social media legislation in larger states.

## Implications

The 2024 election flipped the Senate and Presidency from Democrat to Republican, creating a trifecta in the federal government. This means legislation such as the Kids Online Safety Act (KOSA) and COPPA 2.0 may have a perilous future on the federal level. On one hand, Republicans such as Mike Johnson have indicated opposition to KOSA [@johnson]. On the other hand, Trump appointees, such as incoming surgeon general, have called for banning social media [@napier2024]. It remains unclear how policymakers may proceed with social media regulation on a federal level.

State governments may continue to push for bills to regulate social media. This however may make it hard for technology companies to operate platforms. A state by state system of technology regulation could lead to fragmented laws, creating a complex web of rules for any technology company to operate in.

## Drawbacks and Areas for Future Analysis

The analysis only includes data from the beginning of 2024 through October 2024. Since the beginning of October, states may have passed new legislation changing the status of these legislation. Further, states have passed legislation before 2024. A future area of inquiry may consider the trends in state legislation of children and social media over the past few years.

The data set only analyzed the partisan control of state legislators, not the individual political affiliations of those sponsoring each bill. Future analysis in the political affiliations of those sponsoring the bills may provide better insight into the relationship between partisanship and sponsoring of child safety legislation.

# Conclusion

Federal regulation may prove to be more effective at both creating safer social media platforms, alongside offering a more streamlined system for child safety and wellness online. However, given the unknown stance of the incoming Republican-controlled federal government on social media regulation, state governments may be the only avenue for social media regulation.

{{< pagebreak >}}

# Citations
